**Application of database-independent approach to assess the quality of OTU picking methods**

### 2016-03-26 papers/msystems_2016/Schloss_Commentary_mSystems_2016.Rmd
Editing text

The recent analysis of Kopylova et al [-@Kopylova2016] repeated many of the benchmarking strategies employed by previous researchers. Many algorithm developers have clustered sequences from simulated communities or sequencing data from synthetic communities of cultured organisms and quantify how well the OTU assignments matched the organisms' taxonomy [@Edgar2013; @Mah2014; @Mah2015; @Barriuso2011; @Bonder2012; @Chen2013; @Huse2010; @May2014; @Cai2011; @Sun2011; @White2010; @AlGhalith2016]. Although an OTU definition would ideally match bacterial taxonomy, bacterial taxonomy has proven itself to be fluid and reflect the biases of various research interests. Furthermore, it is unclear how the methods scale to sequences from the novel organisms we are likely to encounter in deep sequencing surveys. In a second approach, developers have compared the time and memory required to cluster sequences in a dataset [@Sun2009; @Cai2011; @Rideout2014; @Mah2014]. These are valid parameters to assess when judging a clustering method, but indicate little regarding the quality of the OTU assignments. For example, reference-based methods are very efficient, but do a poor job of reflecting the genetic diversity within the community when novel sequences are encountered [@Westcott2015]. In a third approach, developers have compared the number of OTUs generated by various methods for a common dataset [@Edgar2013; @Kopylova2016]. Although methods need to guard against excessive splitting of sequences across OTUs, by focusing on minimizing the number of OTUs in a community developers excessively lumping sequences that do not belong together. In a fourth approach a metric of OTU stability has been proposed as a way to assess algorithms [@He2015]. Although it is important that the methods generate reproducible OTU assignments when the initial order of the sequences is randomized, this metric ignores the possibility that the variation in assignments may be equally robust or that the assignments by a highly reproducible algorithm may be quite poor. In a final approach, some developers have assessed the quality of clustering based on the method's ability to generate the same OTUs generated by other methods [@Schmidt2014Limits; @Rideout2014]. Unfortunately, without the ability to ground truth any method, such comparisons are tenuous. There is a need for an objective metric to assess the quality of OTU assignments.

@@

**Figure 1. Comparison of OTU quality generated by multiple algorithms applied to four datasets.** The nearest, average, and furthest neighbor clustering algorithms were used as implemented in mothur (v.1.37)[@Schloss2009]. Abundance (AGC) and Distance-based greedy clustering (DGC) were implemented using USEARCH (v.6.1) and VSEARCH (v.1.5.0)[@He2015; @Edgar2013; @Rognes2015]. Other *de novo* clustering algorithms included Swarm (v.2.1.1)[@Mah2014; @Mah2015], OTUCLUST (v.0.1)[@Albanese2015], and Sumaclust (v.1.0.20). The MCC values for Swarm were determined by selecting the distance threshold that generated the maximum MCC value for each dataset. The USEARCH and SortMeRNA (v.2.0) closed-reference clusterings were performed using QIIME (v.1.9.1) [@Kopylova2012; @Caporaso2010]. Closed-reference clustering was also performed using VSEARCH (v.1.5.0) and NINJA-OPS (v.1.3.2) [@AlGhalith2016]. The order of the sequences in each dataset was randomized thirty times and the intra-method range in MCC values was smaller than the plotting symbol. MCC values were calculated using mothur.

>>>>

The recent analysis of Kopylova et al [-@Kopylova2016] repeated many of the benchmarking strategies employed by previous researchers. Many algorithm developers have clustered sequences from simulated communities or sequencing data from synthetic communities of cultured organisms and quantified how well the OTU assignments matched the organisms' taxonomy [@Edgar2013; @Mah2014; @Mah2015; @Barriuso2011; @Bonder2012; @Chen2013; @Huse2010; @May2014; @Cai2011; @Sun2011; @White2010; @AlGhalith2016]. Although an OTU definition would ideally match bacterial taxonomy, bacterial taxonomy has proven itself to be fluid and reflect the biases of various research interests. Furthermore, it is unclear how the methods scale to sequences from the novel organisms we are likely to encounter in deep sequencing surveys. In a second approach, developers have compared the time and memory required to cluster sequences in a dataset [@Sun2009; @Cai2011; @Rideout2014; @Mah2014]. These are valid parameters to assess when judging a clustering method, but indicate little regarding the quality of the OTU assignments. For example, reference-based methods are very efficient, but do a poor job of reflecting the genetic diversity within the community when novel sequences are encountered [@Westcott2015]. In a third approach, developers have compared the number of OTUs generated by various methods for a common dataset [@Edgar2013; @Kopylova2016]. Although methods need to guard against excessive splitting of sequences across OTUs, by focusing on minimizing the number of OTUs in a community developers risk excessively lumping sequences together that are not similar. In a fourth approach, a metric of OTU stability has been proposed as a way to assess algorithms [@He2015]. Although it is important that the methods generate reproducible OTU assignments when the initial order of the sequences is randomized, this metric ignores the possibility that the variation in assignments may be equally robust or that the assignments by a highly reproducible algorithm may be quite poor. In a final approach, some developers have assessed the quality of clustering based on the method's ability to generate the same OTUs generated by other methods [@Schmidt2014Limits; @Rideout2014]. Unfortunately, without the ability to ground truth any method, such comparisons are tenuous. There is a need for an objective metric to assess the quality of OTU assignments.

@@

**Figure 1. Comparison of OTU quality generated by multiple algorithms applied to four datasets.** The nearest, average, and furthest neighbor clustering algorithms were used as implemented in mothur (v.1.37)[@Schloss2009]. Abundance (AGC) and Distance-based greedy clustering (DGC) were implemented using USEARCH (v.6.1) and VSEARCH (v.1.5.0)[@He2015; @Edgar2013; @Rognes2015]. Other *de novo* clustering algorithms included Swarm (v.2.1.1)[@Mah2014; @Mah2015], OTUCLUST (v.0.1)[@Albanese2015], and Sumaclust (v.1.0.20). The MCC values for Swarm were determined by selecting the distance threshold that generated the maximum MCC value for each dataset. The USEARCH and SortMeRNA (v.2.0) closed-reference clusterings were performed using QIIME (v.1.9.1) [@Kopylova2012; @Caporaso2010]. Closed-reference clustering was also performed using VSEARCH (v.1.5.0) and NINJA-OPS (v.1.5.0) [@AlGhalith2016]. The order of the sequences in each dataset was randomized thirty times and the intra-method range in MCC values was smaller than the plotting symbol. MCC values were calculated using mothur.
